### 2013-05-31 drafts/general_writeup/writeup.tex
Started new introduction to the general writeup document.

\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amsmath}

\usepackage{ amssymb }
\usepackage{listings}
\lstloadlanguages{Haskell}
\lstnewenvironment{code}
    {\lstset{}%
      \csname lst@SetFirstLabel\endcsname}
    {\csname lst@SaveFirstLabel\endcsname}
    \lstset{
      basicstyle=\small\ttfamily,
      flexiblecolumns=false,
      basewidth={0.5em,0.45em},
      literate={+}{{$+$}}1 {/}{{$/$}}1 {*}{{$*$}}1 {=}{{$=$}}1
               {>}{{$>$}}1 {<}{{$<$}}1 {\\}{{$\lambda$}}1
               {\\\\}{{\char`\\\char`\\}}1
               {->}{{$\rightarrow$}}2 {>=}{{$\geq$}}2 {<-}{{$\leftarrow$}}2
               {<=}{{$\leq$}}2 {=>}{{$\Rightarrow$}}2 
               {\ .}{{$\circ$}}2 {\ .\ }{{$\circ$}}2
               {>>}{{>>}}2 {>>=}{{>>=}}2
               {|}{{$\mid$}}1               
    }

@@

\newcommand{\learner}{\emph{Learner }}
\newcommand{\evaluator}{\emph{Evaluator }}

\section{The Learning Problem}
We want to characterize the general task of the learner, whom we will
call Linda. At various intervals, Linda is provided with sensory data
$x$. In reponse, she is given the opportunity of submitting a program
$\phi$ to an evaluator, henceforth named Dr. Eval. Dr. Eval produces a
reward $R(\phi, x; \bar{y})$ which he delivers back to Linda, where
$\bar{y}$ simply denotes any other variables Linda is not aware of
that influences Dr. Eval's reward. Linda processes the information,
hopefully learning from her experience with Dr. Eval.

What kinds of scenarios can this learning paradigm (naturally)
contain? Most basically, Linda might be a simple organism: the data
$x$ might be local chemical gradients, the program $\phi$ the motor
program Linda will execute upon sensing $x$, and Dr. Eval might return
the cumulative effect upon Linda of executing that motor program. Note
that in this case $\phi$ does not take $x$ as an input. It has no
input; it is merely a motor program built anew for each sensory datum
that Linda receives.

Of course, we might want Linda, the still humble organism, to be
capable of learning a higher-level planning procedure. Such a
procedure would be a program that, given sensory data, would produce a
motor program. In our framework, this would be represented by letting
$x$ stand for a list of data (maybe several days worth of discrete
food finding missions) and letting $\phi$ be the planing
procedure. Dr. Eval would accordingly take $\phi$, apply it to each
datum in $x$ and return a cumulative reward of some sort. We will
defer for now the problem of deferred reward, in which most learners
do not receive immediate signals about the utility of their actions.

Moving upward somewhat in the hierarchy of intelligent beings,
consider Linda the electrical engineer. She is requested to construct
a circuit $\phi$ to a certain specification $x$. Dr. Eval, her
professor tests the circuit against the requested specification and
gives her a grade. Or perhaps she doubles as the evaluator. As in the
case of Linda the simple organism, we can conceive of $\phi$ as the
general circuit creation procedure rather than the circuit, and create
a higher-level learning problem.

A different perspective on this learning problem is one that is
removed from this supervised reinforcement paradigm: instead of
acheiving a reward of some sort, we can imagine that Linda wants to
find the program $\phi$ which generated the data $x$. Here, Dr. Eval
is the world or some simulator of the world and a comparator which
signals to Linda whether she has successfully simulated the data. This
paradigm corresponds to a deterministic version of the generative
model learning task common these days in probabilistic machine
learning and AI.

TODO: concept learning
TODO: physicist (or scientist in general)

\section{A Universal Basis for Computation}
Our goal is to explain the minimal conditions necessary for concept
learning, and to do so we posit a minimal model of
computation. Various universal models of computation have been
proposed; the most famous is the Turing machine. Though the Turing
machine is an intuitive abstract machine, and though its construction
corresponds to various parts of real-world computing machines,
formally simpler models of computation have been proposed. Most
importantly is the Lambda calculus, proposed by Church, and the basis
of modern functional programming languages. The Lambda calculus is a
rewrite system that is provably equally powerful to Turing
machines. Despite its popularity, the implementation of the Lambda
calculus is complicated by the presence of variables bound by the
abstracting $\lambda$s. The necessary relationships between variables
and their binding $\lambda$s means that the syntax of the Lambda
calculus is not entirely modular. That is, if we represent an
expression in the lambda calculus as a tree, then every subtree of this
expression tree is not itself a valid expression tree in the langauge.

An alternative to the $\lambda$ calculus we use the variable-free
model of computation known as Combinatory Logic (CL). CL was
introduced by Shonfinkel in order to avoid of the mathematical
complexities of $\lambda$-calculus. The CL has a simpler syntax than
$lambda$-calculus: a combinator $C$ is merely the application of two
other combinators $ C = A B$. The semantics of the language is defined
by the primitive combinators chosen for the particular instance of
$CL$. It is well known that a universal basis of two primitive
combinators exists, the $SK$ basis. However, expressions in CL are
much more readible if we extend this primitive basis with a few other
high order combinators. Namely, we include identity $I$, and two
composition combinators $B$ and $C$.

\section{History and related literature}
Angluin 1988~\cite{angluin1988queries}: investigates the complexity of
learning concepts that are subsets of a hypothesis space via a number
of primitive queries such as membership, equivalence, subset, etc. \\

@@

\section{Domains for concept learning}

\section{Coding Theory}
Consider a distribution $P(\cdot)$ over strings in some set $X \in
\{0, 1\}^*$. Our goal is to find a code $D$ that minimizes the
expected length of a sample from $X$.

@@

\section{Intuitive Psychology}
How can we express the knowledge and relationships of intuitive
psychology as a set of expressions in combinatory logic? To simplify
this task, we will attempt to express these relationships in Haskell,
a sugared and polymorphically typed lambda calculus, knowing that any
expression in this lambda calculus can be systematically translated
into combinatory logic. Most importantly, any modularization and
compositional structure present in the Haskell code will remain
present in its CL translation.

Let's imagine the following situation. You observe several agents
(say, Josh, Max, and Chris) in various scenarios in which they choose,
from among the various options, one of three bars (say, a Zone bar, a
Luna bar and a dark chocolate bar). We will try to represent our
knowledge about these agents, their propensities, and the properties
of these bars in Haskell, as a demonstration that such a
representation is both possible in this langauge, and can be somewhat
natural.

Let's first define the primitive types and combinators at our disposal:

\begin{lstlisting}[mathescape]
-- | Primitive Types
-------------------------
Int -- integers 1, 2, etc.
Bool -- {True, False}
Person $\triangleq$ Int
Object $\triangleq$ Int
Scenario  $\triangleq$ (Person, [Object])

-- | Primitive Combinators
------------------------
-- List Combinators:
map :: (a -> b) -> [b] -> [a]
foldl1 :: (a -> a -> a) -> [a] -> a
elem :: a -> [a] -> Bool
head :: [a] -> a
tail :: [a] -> [a]
case :: [(Bool, a)] -- a set of boolean tests and the result if True
        -> a -- the default result (else clause) 
        -> a
(:) :: a -> [a] -> [a] -- cons
(++) :: [a] -> [a] -> [a] -- concatenate

-- Boolean Combinators
If _ Then _ Else :: Bool -> a -> a
And :: Bool -> Bool -> Bool
Or ::  Bool -> Bool -> Bool
Not ::  Bool -> Bool -> Bool

-- Arithmetic Combinators
(>) :: Int -> Int -> Bool
(<) :: Int -> Int -> Bool
(==) :: Int -> Int -> Bool
(+) ::  Int -> Int -> Bool
(-) :: Int -> Int -> Bool
1, 2, $\dots$, 10 :: Int -- | the integers, 1 through 10

-- Givens
------------

Josh :: Person -- | this is just shorthand, i.e., Josh would be mapped
                  to some integer, say, 1, Max to 2, etc..
Max :: Person
Chris :: Person

ZoneBar :: Object
LunaBar :: Object
ChocolateBar :: Object

-- Learned representation
-------------------------
choose' :: Scenario -> Object
choose' (person, objs) = max' (map (util' person) objs)

max' :: [Int] -> Int
max = foldl1 (>)

util' = Person -> Object -> Int
util' person obj = case (utilityCases' person obj) 0 -- | 0 is the default utility in this case

utilityCases' :: [Person -> Object -> (Bool, Int)]
utilityCases' = \p -> \o -> (
                     \p -> \o -> (isHealthyPerson p, healthyPersonCalculus o)

\end{lstlisting}

>>>>

\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amsmath} 
\usepackage{algpseudocode}
\usepackage{ amssymb }
\usepackage{listings}
\usepackage{syntax}

\newcommand{\funtype}[2]{#1 \rightarrow #2}
\newcommand{\weightedrule}[3]{[#1 \mapsto #2 \, | \, #3]}
\newcommand{\typedexp}[2]{#1 : #2} 
\newcommand{\type}[1]{type(#1)}

\renewcommand{\grammarlabel}[2]{#1 #2}
\newcommand{\given}{\, | \,}

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\lstloadlanguages{Haskell}
\lstnewenvironment{code}
    {\lstset{}%
      \csname lst@SetFirstLabel\endcsname}
    {\csname lst@SaveFirstLabel\endcsname}
    \lstset{
      basicstyle=\small\ttfamily,
      flexiblecolumns=false,
      basewidth={0.5em,0.45em},
      literate={+}{{$+$}}1 {/}{{$/$}}1 {*}{{$*$}}1 {=}{{$=$}}1
               {>}{{$>$}}1 {<}{{$<$}}1 {\\}{{$\lambda$}}1
               {\\\\}{{\char`\\\char`\\}}1
               {->}{{$\rightarrow$}}2 {>=}{{$\geq$}}2 {<-}{{$\leftarrow$}}2
               {<=}{{$\leq$}}2 {=>}{{$\Rightarrow$}}2 
               {\ .}{{$\circ$}}2 {\ .\ }{{$\circ$}}2
               {>>}{{>>}}2 {>>=}{{>>=}}2
               {|}{{$\mid$}}1               
    }

@@

\newcommand{\learner}{\emph{Learner }}
\newcommand{\evaluator}{\emph{Evaluator }}
\section{The problem}
In all but the simplest problems, problem solving agents suffer from vast
search spaces. Much of the research in AI, machine learning, and optimization
focuses on how to overcome this search problem. Local information can help find
better nearby solutions but make few gaurantees about global optima. Heuristic
search techniques make use of the structure of the search space and information
about the utility of partial solutions to hopefully minimize the numeber of
candidate solutions that need to be explored. Other frameworks attempt to
approximate the target search space with a more tractable, convex, surrogate
search space.

But since the begining of AI, a central idea -- at center stage at times and
lurking in the shadows at others -- is that as problem solving agents solve
more problems they should learn to solve problems better.

Here, we present a unified perspective on learning to solve problems, grounding the
problem as one of Bayesian inference over a grammar that generates candidate
problem solutions.

Types:

\begin{grammar}
<$t$> ::= $\tau$, $\sigma$  (type variables)
  \alt $a$, $b$ (concrete types e.g. real, bool, int, etc.)
  \alt $\funtype{t'}{t}$ (function type)
\end{grammar}

Expressions: 
\begin{grammar}
<$\typedexp{e}{t}$> ::= $\typedexp{c}{t'}$ s.t. $t' = mgu(t,
\type{c})$ (substitute a terminal tree that unifies with the requesting type)
  \alt ($\typedexp{e}{\funtype{t'}{t}} \;\;\; \typedexp{e}{t'}$)
\end{grammar}

First we will present the model over solutions in which our problem solving
agent will be doing inference. Second, we will present a hybrid inference and
problem solving algorithm.

\section{The model}
Our goal in this section is to describe a generic probability distribution over
solutions spaces. Learning to search will be implemented by learning a
distribution over the solution space which describes the probability of a 
candidate solution being the solution for a given problem.

Problems vary considerably in their solution spaces. Some are numeric and
continuous, others are integral and discrete, and still others consist of
structured data like trees. Some solutions have a finite number of elements, or
are bounded in some fashion. Others are infinite and unbounded. To present a
framework for general problem solving, we need a representation of a solution
space that subsumes all of these possibilities.

As has long been noted by
Genetic Programming researchers, any solution space can be modeled as a set of
programs that generate space. As an example, if the solution space to a problem
is the set of all boolean circuits constructed from some set of gates, then the
corresponding space of programs is the set of programs that wire up boolean
circuits from those gates.

How does this realization help us? It means that if we can formulate a single
model that places a distribution over solution distributions by describing a
distribution over program generating processes.

We will first describe a simplified version of the hierarchical model over programs that we will be using. We start with a base grammar $\mathcal{G}$ of the form 
\section{The model}
The elements of any solution space can be represented as programs. u
 \section{History and related literature}
Angluin 1988~\cite{angluin1988queries}: investigates the complexity of
learning concepts that are subsets of a hypothesis space via a number
of primitive queries such as membership, equivalence, subset, etc. \\

@@

@@
